A Horrocks-type theorem for even orthogonal $K_2$

### 2023-02-22 dl-paper.tex
Fix typo

Recall that the classical Matsumoto theorem (see~\cite[Theorem~5.10]{Ma69}) allows one to compute the group $\K_2(\Phi, R)$ in the special case when $R=k$ is a field.
Using the modern language of Milnor--Witt K-theory (see~\cite{Mo04}) it can be formulated as follows:
\begin{equation*} \K_2(\Phi, k) = \left\{\begin{array}{ll} \K_2^\mathrm{MW}(k)& \text{if $\Phi$ is symplectic,}\\ \K_2^\mathrm{M}(k) & \text{otherwise.}\end{array}\right. \end{equation*}
In the following lemma we recall the computation of the group $\K_2(\Phi, R)$ in the case $R=k[X, X\inv]$.
\begin{lemma}[Hurrelbrink--Morita--Rehmann]\label{K2-laurent-field} Let $\Phi$ be a reduced irreducible root system and $k$ be an arbitrary field. Then there is a split exact sequence of abelian groups
\[\begin{tikzcd} 0 \arrow{r} & \K_2(\Phi, k) \arrow{r} & \K_2(\Phi, k[X, X^{-1}]) \arrow{r} & H(\Phi, k) \arrow{r} & 0, \end{tikzcd}\text{ in which}\]
The group $H(\Phi, k)$ admits the following description. One has $H(\Phi, k) \cong \K_1^\mathrm{M}(k) \cong k^\times$ if $\Phi$ is nonsymplectic. If $\Phi$ is symplectic and $\mathrm{char}(k)\neq 2$ then $H(\Phi, k) \cong \K_1^\mathrm{MW}(k)$. \end{lemma}
\begin{proof} Let us first consider the case of nonsymplectic $\Phi$, in which one can find a long root $\alpha\in \Phi$ in such a way that there is a commutative diagram of abelian groups
\[\begin{tikzcd}[column sep=5pc] k^\times \arrow{drr}[swap]{\{-, X\}} \arrow{r}{h}  & \K_2(\Phi, k[X, X^{-1}]) \arrow{r} & \K_2(\Phi, k(X)) \arrow{d}{\cong} \\
 & & \K_2^\mathrm{M}(k(X)), \end{tikzcd}\]
in which $h = \{ -, X \}_{\alpha}$ and the vertical map is an isomorphism by Matsumoto's theorem.
Notice that the diagonal map is split by the obvious residue homomorphism and therefore is injective. This, in turn, implies that $h$ is also injective.
For $\Phi \neq \rG_2$ the assertion of the lemma follows from~\cite[Satz~3]{Hur77} which asserts that for a nonsymplectic $\Phi\neq\rG_2$ one has $\K_2(\Phi, k[X, X\inv]) = \mathrm{Im}(h) \oplus \K_2(\Phi, k)$.
To obtain the assertion in the case $\Phi = \rG_2$ repeat the proof of~\cite[Satz~3]{Hur77} using \cite[Corollary~6]{AM88} instead \cite[Korollar~4]{Hur77}.

>>>>

Recall that the classical Matsumoto theorem (see~\cite[Theorem~5.10]{Ma69}) allows one to compute the group $\K_2(\Phi, R)$ in the special case when $R=k$ is a field.
Using the modern language of Milnor--Witt K-theory (see~\cite{Mo04}) it can be formulated as follows:
\begin{equation*} \K_2(\Phi, k) = \left\{\begin{array}{ll} \K_2^\mathrm{MW}(k)& \text{if $\Phi$ is symplectic,}\\ \K_2^\mathrm{M}(k) & \text{otherwise.}\end{array}\right. \end{equation*}
In the following lemma we recall the computation of the group $\K_2(\Phi, R)$ in the case $R=k[X, X\inv]$.
\begin{lemma}[Hurrelbrink--Morita--Rehmann]\label{K2-laurent-field} Let $\Phi$ be a reduced irreducible root system and $k$ be an arbitrary field. Then there is a split exact sequence of abelian groups
\[\begin{tikzcd} 0 \arrow{r} & \K_2(\Phi, k) \arrow{r} & \K_2(\Phi, k[X, X^{-1}]) \arrow{r} & H(\Phi, k) \arrow{r} & 0, \end{tikzcd}\text{ in which}\]
The group $H(\Phi, k)$ admits the following description. One has $H(\Phi, k) \cong \K_1^\mathrm{M}(k) \cong k^\times$ if $\Phi$ is nonsymplectic. If $\Phi$ is symplectic and $\mathrm{char}(k)\neq 2$ then $H(\Phi, k) \cong \K_1^\mathrm{MW}(k)$. \end{lemma}
\begin{proof} Let us first consider the case of nonsymplectic $\Phi$, in which one can find a long root $\alpha\in \Phi$ in such a way that there is a commutative diagram of abelian groups
\[\begin{tikzcd}[column sep=5pc] k^\times \arrow{drr}[swap]{\{-, X\}} \arrow{r}{h}  & \K_2(\Phi, k[X, X^{-1}]) \arrow{r} & \K_2(\Phi, k(X)) \arrow{d}{\cong} \\
 & & \K_2^\mathrm{M}(k(X)), \end{tikzcd}\]
in which $h = \{ -, X \}_{\alpha}$ and the vertical map is an isomorphism by Matsumoto's theorem.
Notice that the diagonal map is split by the obvious residue homomorphism and therefore is injective. This, in turn, implies that $h$ is also injective.
For $\Phi \neq \rG_2$ the assertion of the lemma follows from~\cite[Satz~3]{Hur77} which asserts that for a nonsymplectic $\Phi\neq\rG_2$ one has $\K_2(\Phi, k[X, X\inv]) = \mathrm{Im}(h) \oplus \K_2(\Phi, k)$.
To obtain the assertion in the case $\Phi = \rG_2$ repeat the proof of~\cite[Satz~3]{Hur77} using \cite[Corollary~6]{AM88} instead of \cite[Korollar~4]{Hur77}.
